\chapter*{Acknowledgements}

There are a great many people I would like to thank for helping me achieve success at Carleton.

My supervisors, Anil Maheshwari and Michiel Smid, have had their doors open to me for my entire tenure at Carleton, ever since my first summer in the Computational Geometry lab way back at the beginning of my undergraduate degree. 
They have endured countless drop-in questions, and provided many impromptu lectures on the walk to lunch any time I needed help navigating a problem.

Much credit is also due to Subhas Nandy of the Indian Statistical Institute, Kolkata, who got me started with this thesis work. 
Thank you for all your hospitality in India, and of course, for all the chai.

The Computational Geometry lab has been an amazing environment to work and learn in. 
The entire faculty of the CG lab, Prosenjit Bose, Anil Maheshwari, Pat Morin, and Michiel Smid, put in countless hours supporting each and every student of the lab. 
I would like to thank Jean-Lou De Carufel for teaching me the art of inequalities, and my fellow students Nima Hoda and Simon Pratt for teaching me as much about our coursework as the lectures did.

I gratefully acknowledge Carleton University, the School of Computer Science, and the National Sciences and Engineering Research Council of Canada for their generous funding support.

Finally, to my amazing wife, Kathryn Bint. 
I would never have survived my degree without your love and unerring support through uncountably many busy evenings and weekends.
