\chapter{Conclusion}
\label{:conclusion}

In this final chapter, we summarize the contributions from the previous chapters as well as possible directions for future research.

\section{Summary of Contributions}
\label{:conclusion:contributions}

This thesis presents several contributions in the area of majority range searching.

In Chapter~\ref{:rectangles}, we developed methods for querying line segments with axis-parallel rectangles. 
At the first, the segments were also axis-parallel and we gave a method for expressing the majority property as an orthogonal range query. 
Next we considered arbitrarily-oriented line segments, and we saw that the uncertainty of where a segment may intersect the query significantly increases the complexity of testing majority expressions.

In Chapter~\ref{:slabs}, we consider axis-parallel line segments against a query region which is either an arbitrarily-oriented slab, or the intersection of two such slabs, which is a generalization of an arbitrarily-oriented rectangle. We gave a method for expressing the majority property as a half-plane query in a dual-space.

In Chapter~\ref{:convexp}, we moved away from line segments to consider convex polygons and majority range queries involving area. 
We developed a method for preprocessing a convex polygon such that the area to one side of any chord can be found in logarithmic time with respect to the number of edges in the polygon. 
We gave an algorithm for decomposing the intersection of a rectangle with a convex polygon in such a way that the area of their intersection could be found in logarithmic time. 
Deciding on majority enclosure is then straight-forward.

In Chapter~\ref{:monotonep}, we considered monotone polygons. 
We gave an algorithm which produces a list of area formulas that we can use to calculate the area of the polygon underneath a query line.
We combined this approach with our preprocessing idea from the convex polygon problem and developed a method to calculate the area of a monotone polygon within a query rectangle in logarithmic time with respect to the number vertices in the polygon. 
With the intersection area calculated, deciding on the majority enclosure property is straight-forward in both this and the convex polygon case.


\section{Future Work}
\label{:conclusion:open-problems}

In this section, we summarize the future work we have discussed in previous chapters.

\begin{enumerate}
\item From Chapter~\ref{:rectangles}, can we remove the limitation that $1/2 < \rho$ when querying on arbitrarily-oriented line segments?  This limitation stems from our current solution relying on the centrepoint of the located segments residing inside the query region.

\item From Chapter~\ref{:rectangles}, can we classify the endpoints of the segments in a way that reduces the number of majority expressions that we need to test.  The correct majority expression depends on which boundaries of the query rectangle are crossed by a segment, but our current classification technique is limited in how well it can identify these.

\item From Chapter~\ref{:rectangles}, in a practical setting, when we are better off saving a constant factor of space by sharing the endpoint classification structures versus saving a $\log{n}$ factor by performing the half-plane queries first.

\item From Chapter~\ref{:slabs}, extending our method to arbitrarily-oriented segments results in a very case-heavy data structure. Moreover, the expressions for testing the majority properties involve several independent combinations of query variables resulting in high-degree half-plane queries. Is there a better method of classification that can reduce this?

\item From Chapter~\ref{:monotonep}, can we modify our approach to building a multi-region formula tree to support querying on simple polygons?  Our tree relies on clean boundaries between regions so that we can query a single formula for any horizontal line query. In a simple polygon setting, two regions may be located one over top of the other without sharing common boundaries.  Attempting to partition the polygon by the boundary of one region can create subregions which break our area formulas.

\end{enumerate}
